On the presentation of orthogonal groups over polynomial rings

### 2019-06-09 dl-paper.tex
Minor fix

For $n\geq 1$ we define ``truncated'' Steinberg groups $\St^{\leq n}(\Phi, B)$ by the set of homogeneous Steinberg generators of degree $\leq n$ modulo Steinberg relations of degree $\leq n$.
The following lemma asserts that most of the relators $\text{R3}^\bot_{d,e}$ in this presentation of $\St^{\leq n}(\Phi, B)$ can be omitted.
\begin{lemma}\label{superfluous-relations}
 For every simply-laced root system $\Phi$ of rank $\geq 3$ and every $n \geq 1$ one can exclude from the presentation of $\St^{\leq n}(\Phi, B)$ all the relators $\text{R3}_{d,e}^\bot$ for which $\max(0,d) + \max(0,e) > 1$.
\end{lemma}
\begin{proof}
For the proof we will need the following commutator identities:
\begin{align} \label{eq:H1ii} [xy, z] = {}^x[y, z] \cdot [x,z];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz].& \end{align}
Without loss of generality we may assume $e \geq d$ and $e > 0$. Let $\alpha$ and $\gamma$ be a pair of orthogonal roots of $\Phi$.
We can find $\beta \in \Phi$ forming an obtuse angle with both $\alpha$ and $\gamma$ (see e.\,g.~\cite[Lemma~3.1.2]{RS76}).

@@

\item In the case $d \leq 0 \leq e \leq n$ the relation $\text{R3}^\bot_{d,e}$ is a consequence of $\text{R3}^\bot_{1,d+e-1}$:
\begin{align*} [x_\alpha(at^d), x_{\gamma}(bt^{e})] = [x_\alpha(at^d), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$ } \\
 = [[x_\alpha(at^d), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{d,1}$ } \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(ab't^{e-1}t^d), x_{-\beta}(t)] & \text{ by $\text{R2}_{d,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$  } \\
 = 1 & \text{ by $\text{R3}^\bot_{d+e-1,1}$.} \end{align*} \end{enumerate}

To see that $\text{R3}^\bot_{d,e}$ is superfluous one must iteratively apply one of the above observations.
Notice that the value $\max(0,d,e)$ decreases strictly whenever $e > 1$. 
Finally, by the first observation $\text{R3}^\bot_{1,1}$ follows from $\text{R3}^\bot_{0,0}$.
\end{proof}

\begin{lemma} For $\Phi=\rA_{\geq 4}, \rD_{\geq 5}, \rE_{6,7,8}$ and $n \geq 1$ the map $i_\Phi\colon \St^{\leq n}(\Phi, B) \to \St^{\leq n+1}(\Phi, B)$ induced by the natural embedding of generators, is an isomorphism. In particular, the natural map $\St^{\leq 1}(\Phi, B) \to \St(\Phi, B)$ is an isomorphism. \end{lemma}
\begin{proof}
 The special case $\Phi=\rA_{\geq 4}$ of the assertion is due to Tulenbayev, see~\cite[Lemma~3.3]{Tu83}. Thus, it remains to deduce from it the assertion in the cases $\Phi=\rD_\ell,\rE_\ell$.

 Denote by $A_4(\Phi)$ the set consisting of all root subsystems of $\Phi$ of type $\rA_4$ and by $G_n$ the free product of groups $\coprod\limits_{\Psi \in A_4(\Phi)} \St^{\leq n}(\Psi, B)$ amalgamated over relations $x^{\Psi_1}_\alpha(a) = x^{\Psi_2}_\alpha(a)$ for all $a \in B_d$, $d \leq n$, $\alpha \in \Psi_1\cap\Psi_2$, $\Psi_1, \Psi_2 \in A_4(\Phi)$.

>>>>

For $n\geq 1$ we define ``truncated'' Steinberg groups $\St^{\leq n}(\Phi, B)$ by the set of homogeneous Steinberg generators of degree $\leq n$ modulo Steinberg relations of degree $\leq n$.
The following lemma asserts that most of the relators $\text{R3}^\bot_{d,e}$ in this presentation of $\St^{\leq n}(\Phi, B)$ can be omitted.
\begin{lemma}\label{superfluous-relations}
 For every simply-laced root system $\Phi$ of rank $\geq 3$ and every $n \geq 1$ one can exclude from the presentation of $\St^{\leq n}(\Phi, B)$ the relators $\text{R3}_{d,e}^\bot$ for which $\max(0,d) + \max(0,e) > 1$.
\end{lemma}
\begin{proof}
For the proof we will need the following commutator identities:
\begin{align} \label{eq:H1ii} [xy, z] = {}^x[y, z] \cdot [x,z];&\\ %= [a,[b,c]]\cdot [b,c] \cdot [a,c];&\\
 \label{eq:H1iii} [x,z] = 1 \text{ implies } [x, [y,z]] = [[x,y],{}^yz].& \end{align}
Without loss of generality we may assume $e \geq d$ and $e > 0$. Let $\alpha$ and $\gamma$ be a pair of orthogonal roots of $\Phi$.
We can find $\beta \in \Phi$ forming an obtuse angle with both $\alpha$ and $\gamma$ (see e.\,g.~\cite[Lemma~3.1.2]{RS76}).

@@

\item In the case $d \leq 0 \leq e \leq n$ the relation $\text{R3}^\bot_{d,e}$ is a consequence of $\text{R3}^\bot_{1,d+e-1}$:
\begin{align*} [x_\alpha(at^d), x_{\gamma}(bt^{e})] = [x_\alpha(at^d), [x_{\beta+\gamma}(b't^{e-1}), x_{-\beta}(t)]] & \text{ by~$\text{R2}_{e-1,1}$ } \\
 = [[x_\alpha(at^d), x_{\beta+\gamma}(b't^{e-1})], {}^{x_{\beta+\gamma}(b't^{e-1})}\!x_{-\beta}(t)] & \text{ by \eqref{eq:H1iii} and $\text{R3}^\angle_{d,1}$ } \\
 = {}^{x_{\beta+\gamma}(b't^{e-1})}\![x_{\alpha+\beta+\gamma}(ab't^{e-1}t^d), x_{-\beta}(t)] & \text{ by $\text{R2}_{d,e-1}$ and $\text{R3}^\angle_{e-1,e-1}$  } \\
 = 1 & \text{ by $\text{R3}^\bot_{d+e-1,1}$.} \end{align*} \end{enumerate}                                                           
Together the above observation show that if the degree of the relator $\text{R3}^\bot_{d,e}$ is $\geq 2$ then it is a consequence of some relator $\text{R3}^\bot_{d',e'}$ of strictly smaller degree.
Finally, by the first observation, $\text{R3}^\bot_{1,1}$ is a consequence of $\text{R3}^\bot_{0,0}$.
\end{proof}

\begin{lemma} For $\Phi=\rA_{\geq 4}, \rD_{\geq 5}, \rE_{6,7,8}$ and $n \geq 1$ the map $i_n\colon \St^{\leq n}(\Phi, B) \to \St^{\leq n+1}(\Phi, B)$ induced by the natural embedding of generators, is an isomorphism. In particular, the natural map $\St^{\leq 1}(\Phi, B) \to \St(\Phi, B)$ is an isomorphism. \end{lemma}
\begin{proof}
 The special case $\Phi=\rA_{\geq 4}$ of the assertion is due to Tulenbayev, see~\cite[Lemma~3.3]{Tu83}. 
 Therefore it can be used to deduce our assertion in the cases $\Phi=\rD_\ell,\rE_\ell$.

 %Denote by $A_4(\Phi)$ the set consisting of all root subsystems of $\Phi$ of type $\rA_4$ and by $G_n$ the free product of groups $\coprod\limits_{\Psi \in A_4(\Phi)} \St^{\leq n}(\Psi, B)$ amalgamated over relations $x^{\Psi_1}_\alpha(a) = x^{\Psi_2}_\alpha(a)$ for all $a \in B_d$, $d \leq n$, $\alpha \in \Psi_1\cap\Psi_2$, $\Psi_1, \Psi_2 \in A_4(\Phi)$.
